---
title: "ESM 244 Lab 1"
author: "Sara Orofino"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Objectives**:

- basic review of data wrangling
- intro to spatial data in R with sf
- intro to shiny apps 

**Packages**: 

```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(tmap)
```

####1. Review: reading, exploring, wrangling, and visualizing data 

a. Read in the data   
```{r, data}

wine <- read_csv("wine_data.csv")

```

b. Data wrangling:
   Select - country, province, winery, region_1, points, price
   Rename - province to state
   Filter fo only California, Oregon, and Washington
   Add a column for price/points
   Arrange price/point from highest to lowest
```{r, wrangling}

wine_new <- wine %>% 
  select(country, province, winery, region_1, points, price) %>% 
  rename(state = province) %>% 
  filter(state == "California" | state == "Oregon" | state == "Washington") %>% 
  mutate(ppd = points/price) %>% 
  arrange(-ppd)
```

c. Summarize data:
   Summarize groups based on state using group_by()
   Find mean, maximum, minimum, and count
   
```{r, summary}

wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE),
    max_ppd = max(ppd, na.rm = TRUE),
    min_ppd = min(ppd, na.rm = TRUE),
    count = length(state)
  )
```

d. A few gg-graphics 
   1. Histograms of price per dollar value by state (side-by-side histograms)
   
```{r, wine histogram}
# number of bins 
k <- 2*((NROW(na.omit(wine_new$ppd)))^(1/3))
        
wine_hist <- ggplot(wine_new, aes(x = ppd)) + 
  geom_histogram(bins = k, aes(fill=state)) +
  facet_wrap(~state, scales = "free") 

wine_hist
```

  2. Scatterplot graph of price vs. points, with color determined by state, adding a linear model trendline on top of it

```{r, scatterplot}

wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color = state), alpha = 0.5) +
  geom_smooth(color = "black", method = "lm", se = FALSE) +
  scale_color_manual(breaks = c("California", "Oregon", "Washington"), values = c("blue", "coral", "purple")) +
  facet_wrap(~state) +
  theme_light()

wine_points
```

